My title

### 2024-11-04 paper/paper.qmd
overview finished

# Plotting the support percentages side-by-side for each state
ggplot(plot_data, aes(x = state, y = support_percentage, fill = candidate)) +
  geom_bar(stat = "identity", position = position_dodge(width = 0.7)) +
  labs(
    title = "Support Percentages for Harris and Trump by State",
    x = "State",
    y = "Support Percentage"
  ) +
  scale_fill_manual(values = c("Harris" = "blue", "Trump" = "red"), name = "Candidate") +
  theme_minimal() +
  theme(axis.text.x = element_text(angle = 90, hjust = 1))

ggplot(harrisdata, aes(x = reorder(state, pct, median), y = pct, fill = state)) +
  geom_boxplot() +
  coord_flip() +  # Flip for better readability
  labs(x = "State", y = "Support Percentage") +
  theme_minimal() +
  theme(legend.position = "none")
```
We focus on the swing states of the U.S, according to  [cite NYT PAPER]

These are the statistics of each swing state by percetnage by state.

```{r}
#| echo: false
#| warning: false
#| message: false
#| label: fig-pollingharris
#| fig-cap: "Polling for Harris By State"

#### Plot data ####
base_plot <- ggplot(harrisdata, aes(x = end_date, y = pct)) +
  theme_classic() +
  labs(y = "Harris percent", x = "Date") +
    theme(legend.position = "bottom")

base_plot +
  geom_point(aes(color = state)) +
  geom_smooth() +
  theme(legend.position = "bottom")

ggplot(harrisdata, aes(x = end_date, y = pct, color = state)) +
  geom_point() +
  geom_smooth(se = TRUE) +
  facet_wrap(~ state, scales = "free_y") +
  labs(y = "Harris Percent", x = "Date", title = "Polling Trends for Kamala Harris by State") +
  theme_minimal()

@@

# Calculate the average predicted support by state
predicted_support_table <- harrisdata %>%
  group_by(state) %>%
  summarize(average_predicted_support = mean(median_prob) * 100)

@@

Arizona is a very important swing state especially as the state has switches which party they voted for in in the last two elections with Trump winning 49% of the vote and Clinton with 45.5%, which considering Arizona is a swing state is a rather significant6 margin. While in the 2020 election Biden ended up winning Arizona with 49.4% of the vote and Trump 49.1%, which is a much smaller margin of victory and what makes Arizona so questionable in the 2024 election. We can see in @fig-pollingharris that while the polling numbers for Harris started somewhat low around 42%, the have gone up significantly and have continued to trend upwards in the most recent polls. Although, in the @fig-supportstate trump still has larger support in the state, and the largest polling difference in Trump's favour out of all the swing states polled.

We believe Arizona will vote: **IDK**

@@

Considering the last two elections Georgia is very similar to Arizona, as in 2016 Trump won the state with 50.8% to Clinton's 45.6%. While in the 2020 election Biden won with a slim victory of 49.5% to Trump's 49.3%, which again is very similar to Arizona's last two election results with significant republican victory in 2016 and a rather marginal win in 2020, which was the first democratic victory for the state since 1992. Georgia is in an interesting situation as lot of the democratic vote in the state comes from the black community, although Harris a has less support from male black population, which could very realistically result in a Trump victory in the state in the upcomming election.

>>>>

# Plotting the support percentages side-by-side for each state
ggplot(plot_data, aes(x = state, y = support_percentage, fill = candidate)) +
  geom_bar(stat = "identity", position = position_dodge(width = 0.9)) +
  labs(
    title = "Support Percentages for Harris and Trump by State",
    x = "State",
    y = "Support Percentage"
  ) +
  scale_fill_manual(values = c("Harris" = "blue", "Trump" = "red"), name = "Candidate") +
  theme_minimal() +
  theme(axis.text.x = element_text(angle = 90, hjust = 1))
```

In this analysis, we examine polling trends for Kamala Harris across key swing states, focusing on recent polling data from Arizona, Georgia, Michigan, Nevada, North Carolina, Pennsylvania, and Wisconsin [CITE NYT PAPER]. As shown in @fig-supportstate, which displays the support percentages for both Harris and Donald Trump side-by-side by state, the data reveals a competitive race, with relatively close support levels for both candidates in most states. While Trump has a slight edge in Arizona, Georgia, and North Carolina, Harris maintains a modest lead in Michigan, Nevada, and Pennsylvania.

Moving to @fig-supportharris, which isolates Harris's support percentages across states, we gain insight into the consistency and variability of her support. The box plot highlights states like Wisconsin and Pennsylvania, where the smaller interquartile ranges indicate less variability, suggesting that Harris's support is more stable in these regions. Conversely, states such as Nevada and Georgia show greater variability, as reflected in the larger spread of the boxes and whiskers, indicating that Harris's support fluctuates more in these areas.

```{r, fig.height= 3, fig.width= 6}
#| echo: false
#| warning: false
#| message: false
#| label: fig-supportharris
#| fig-cap:  "Support Percentage for Harris by State"

ggplot(harrisdata, aes(x = reorder(state, pct, median), y = pct, fill = state)) +
  geom_boxplot() +
  coord_flip() +  # Flip for better readability
  labs(x = "State", y = "Support Percentage") +
  theme_minimal() +
  theme(legend.position = "none")
```

To explore temporal trends, we introduce the variable `end_date_num`, representing the number of days since Harris declared her candidacy. The next figures examine support trends over time in individual states.

```{r, fig.width= 6, fig.height= 3}
#| echo: false
#| warning: false
#| message: false
#| label: fig-supportWP
#| fig-cap:  "Support Percentage for Harris by State since Harris declared as Candidate for Wisconsin and Pennsylvania"

# Filter for a subset of states to plot separately
subset1 <- harrisdata %>% filter(state %in% c("Wisconsin", "Pennsylvania"))

# Plot the first subset
ggplot(subset1, aes(x = end_date, y = pct, color = state)) +
  geom_point() +
  geom_smooth(se = TRUE) +
  facet_wrap(~ state, scales = "free_y") +
  labs(y = "Harris Percent", x = "Date") +
  theme_minimal() +
    scale_x_date(
    date_breaks = "1 month",            # Set major breaks at monthly intervals
    date_minor_breaks = "1 week",       # Optional: add minor weekly breaks for granularity
    date_labels = "%b %d"               # Format to show abbreviated month and day (e.g., "Aug 01")
  ) +
   theme(legend.position = "none")

```
In @fig-supportWP, Pennsylvania and Wisconsin display distinct trends. In Pennsylvania, support for Harris has been relatively consistent since early September, stabilizing around 48.5%. In contrast, Wisconsin shows a peak in support in mid-September, followed by a gradual decline that stabilizes in October. These observations suggest that while Pennsylvania remains a consistent stronghold for Harris, Wisconsin’s voter sentiment may be more susceptible to fluctuations.

```{r, fig.width= 6, fig.height= 3}
#| echo: false
#| warning: false
#| message: false
#| label: fig-supportAM
#| fig-cap:  "Support Percentage for Harris by State since Harris declared as Candidate"

# Filter for the remaining states
subset2 <- harrisdata %>% filter(state %in% c( "Arizona", "Michigan"))

# Plot the second subset
ggplot(subset2, aes(x = end_date, y = pct, color = state)) +
  geom_point() +
  geom_smooth(se = TRUE) +
  facet_wrap(~ state, scales = "free_y") +
  labs(y = "Harris Percent", x = "Date") +
  theme_minimal() +
      scale_x_date(
    date_breaks = "1 month",
    date_minor_breaks = "1 week",
    date_labels = "%b %d" 
  ) +
   theme(legend.position = "none")
```
@fig-supportAM further examines Harris's support trends in Arizona and Michigan. Both states exhibit an upward trend in support starting in early October, which may suggest increasing favorability or strategic campaign efforts in these regions. This upward trajectory could indicate an opportunity for Harris to solidify her support in these competitive areas.

```{r, fig.width= 7, fig.height= 3}
#| echo: false
#| warning: false
#| message: false
#| label: fig-supportGNN
#| fig-cap:  "Support Percentage for Harris by State since Harris declared as Candidate"

subset3 <- harrisdata %>% filter(state %in% c("Georgia", "North Carolina", "Nevada"))

# Plot the first subset
ggplot(subset3, aes(x = end_date, y = pct, color = state)) +
  geom_point() +
  geom_smooth(se = TRUE) +
  facet_wrap(~ state, scales = "free_y") +
  labs(y = "Harris Percent", x = "Date") +
  theme_minimal() +
        scale_x_date(
    date_breaks = "1 month",
    date_minor_breaks = "1 week",
    date_labels = "%b %d" 
  ) +
   theme(legend.position = "none")
```
In @fig-supportGNN, we analyze Georgia, Nevada, and North Carolina. Nevada demonstrates the highest variability in Harris's support, which could reflect the state’s dynamic political response to recent campaign efforts. Georgia shows a steady upward trend, indicating growing support for Harris, while North Carolina reveals a slight decline since October, suggesting a possible shift in voter support.

Overall, these visualizations illustrate that voter support for Harris varies significantly across states and over time. Notably, many states exhibit shifts in mid-September, indicating that voter opinions are sensitive to external factors or campaign dynamics during this period. This temporal variable, alongside state-specific characteristics, will inform the predictive model to better understand and forecast support for Harris in key regions.

@@

# Calculate the average predicted support by state
predicted_support_table <- harrisdata %>%
  group_by(state) %>%
  summarize(average_predicted_support = round(mean(median_prob) * 100,2))

@@

If my paper were 10 pages, then should be be at least 2.5 pages. The discussion is a chance to show off what you know and what you learnt from all this.

@@

Please don't use these as sub-heading labels - change them to be what your point actually is.
